Compositional equivalences based on Open pNets

### 2021-09-16 2016-Weak-FH-Bisim/WeakBisim.tex
Update WeakBisim.tex
Quelques typos pages 1 et 2

Establishing equivalences between programs or system is crucial both for verifying correctness of programs, by establishing that two implementations are equivalent, and for justifying optimisations and program transformations, by establishing that a modified program is equivalent to the source one. There exist several equivalence relations for programs, and bisimulations are among the most versatile of these equivalences. Among bisimulation relations one distinguishes strong bisimulation, that requires that each action of a program is simulated by a single action of the equivalent program, a weak bisimulation that  is a coarser relations, allowing some of the actions to be
  invisible or internal moves, and thus not simulated by the equivalent program.

@@

In the nineties, several 
works extended the basic behavioural models based on labelled
transition systems to address value-passing or parameterised systems, using
various symbolic encodings of the
transitions~\cite{deSimone85,Larsen87}. 
These works use the term \emph{parameter} to designate systems where variables defining the system as a strong influence on their behaviour: in parametrised systems, parameters can typically influence the number of processes in the system or the way they interact.
In \cite{IngolfsdottirL:2001,HennessyLin:TCS95}, Lin, Ingolfsdottir and Hennessy developed a full hierarchy of
bisimulation equivalences, together with a proof system, for value passing CCS,
including notions of symbolic behavioural semantics and various symbolic 
bisimulations (early and late, strong and weak, and their congruent versions). They also extended this work to models with explicit assignments \cite{Linconcur96}.
Separately J. Rathke~\cite{HennessyRathke:TCS98} defined another symbolic semantics for 
a parameterised broadcast calculus, together with strong and weak bisimulation
equivalences, and developed a symbolic model-checker based on a tableau
method for these processes. 30 years later, no
practical verification approach and no verification platform are
using this kind of approaches to provide proof methods for
value-passing processes or open process expressions.

@@

\subsubsection*{pNets and open automata}
pNets is a convenient model to model concurrent systems in a hierarchical and parametrised way. The coordination between processes is expressed as synchronisation vectors that allows the definition of complex and expressive synchronisation patterns.
open pNets are pNets for which one of the element in the hierarchy is still undefined, such undefined element is called \emph{hole}; a hole can be \emph{filled} later by providing another pNet.
 The semantics of pNets can be expressed as a translation to a labelled transition system, but only if the pNet has no parameter or it has holes. Adding parameters to a LTS is quite standard but enabling holes inside LTSs is not a well-defined notion. 
We thus define \emph{open automata} that can be seen as LTSs with parameters and holes. The transitions of open automata are much more complex than transitions of an LTS as the firing on a transition depends on parameters and actions that are symbolic. This article defines the notion of \emph{open transition} a transition that is symbolic in terms of parameters and coordinated actions.

Contrarily to pNets, open automata are not hierarchical structure, thus they are more convenient for formal reasoning but not adapted to the definition of a complex and structured system like pNets. Additionally, open transitions are expressed in terms of logics more than in terms of synchronised actions, synchronisation vectors on the contrary make it easy to express synchronisations that exist in process algebra or distributed systems.

This article defines pNets and illustrate with an example how they can be used to provide the model of a communicating system. Then we introduce open automata to provide a semantics to open pNets and introduce bisimulation relations and their properties.
Open automata can be also seen as an algebra that can be studied independently from its application to pNets but the composition properties make more sense in a hierarchical model like pNets.

>>>>

Establishing equivalences between programs or systems is crucial both for verifying correctness of programs, by establishing that two implementations are equivalent, and for justifying optimisations and program transformations, by establishing that a modified program is equivalent to the source one. There exist several equivalence relations for programs, and bisimulations are among the most versatile of these equivalences. Among bisimulation relations one distinguishes strong bisimulation, that requires that each action of a program is simulated by a single action of the equivalent program, a weak bisimulation that  is a coarser relation, allowing some of the actions to be
  invisible or internal moves, and thus not simulated by the equivalent program.

@@

In the nineties, several 
works extended the basic behavioural models based on labelled
transition systems to address value-passing or parameterised systems, using
various symbolic encodings of the
transitions~\cite{deSimone85,Larsen87}. 
These works use the term \emph{parameter} to designate systems where variables defining the system have a strong influence on their behaviour: in parametrised systems, parameters can typically influence the number of processes in the system or the way they interact.
In \cite{IngolfsdottirL:2001,HennessyLin:TCS95}, Lin, Ingolfsdottir and Hennessy developed a full hierarchy of
bisimulation equivalences, together with a proof system, for value passing CCS,
including notions of symbolic behavioural semantics and various symbolic 
bisimulations (early and late, strong and weak, and their congruent versions). They also extended this work to models with explicit assignments \cite{Linconcur96}.
Separately J. Rathke~\cite{HennessyRathke:TCS98} defined another symbolic semantics for 
a parameterised broadcast calculus, together with strong and weak bisimulation
equivalences, and developed a symbolic model-checker based on a tableau
method for these processes. 30 years later, no
practical verification approach and no verification platform are
using this kind of approaches to provide proof methods for
value-passing processes or open process expressions.

@@

\subsubsection*{pNets and open automata}
pNets is a convenient model to model concurrent systems in a hierarchical and parametrised way. The coordination between processes is expressed as synchronisation vectors that allow the definition of complex and expressive synchronisation patterns.
Open pNets are pNets for which some elements in the hierarchy are still undefined, such undefined elements are called \emph{holes}; a hole can be \emph{filled} later by providing another pNet.
 The semantics of pNets can be expressed as a translation to a labelled transition system, but only if the pNet has no parameter and no holes. Adding parameters to a LTS is quite standard but enabling holes inside LTSs is not a well-defined notion. 
We thus define \emph{open automata} that can be seen as LTSs with parameters and holes. The transitions of open automata are much more complex than transitions of an LTS as the firing on a transition depends on parameters and actions that are symbolic. This article defines the notion of \emph{open transition} a transition that is symbolic in terms of parameters and coordinated actions.

Contrarily to pNets, open automata are not hierarchical structures, thus they are more convenient for formal reasoning but not adapted to the definition of a complex and structured system like pNets. Additionally, open transitions are expressed in terms of logics more than in terms of synchronised actions, synchronisation vectors on the contrary make it easy to express synchronisations that exist in process algebra orin specification and high-level languages for distributed systems.

This article defines pNets and illustrates with an example how they can be used to provide the model of a communicating system. Then we introduce open automata to provide a semantics to open pNets and introduce bisimulation relations and their properties.
Open automata can be also seen as an algebra that can be studied independently from its application to pNets but their composition properties make more sense in a hierarchical model like pNets.
